opXRD: Open Experimental Powder X-ray Diffraction Database

### 2024-06-09 chapters/01_introduction.tex
Commited first paragraph in "Our dataset" section from external author

% On Rietveld refinement
After fabricating a new material, a number of analysis techniques can be used to characterize the sample.
The most important method to determine the crystal structure of a new material is powder X-ray diffraction. 
To analyze pXRD data, an initial guess is optimized using Rietveld refinement. In this process, an initial crystal structure including unit cell parameters is provided by the operator of the analysis software and fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of expected pXRD diffractograms given the current crystal structure.\cite{Dinnebier2019, Cano2021}
As Rietveld refinement is a local optimization method, the result of the refinement procedure is generally only as good as the initial guess it was provided with.
Therefore, it strongly relies on the experience of the operator to provide an accurate initial guess of the crystal structure.

% Manual vs (ML) automated Refinement
Employing manually operated Rietveld refinement, including determining a good initial structure for the optimization, is not without issues.
It is certainly not scalable to the degree required to keep up with advances in throughput and efficiency in other steps of the experimentation pipeline.
Automation of the Rietveld refinement requires the automation of finding initial guess structures, which is highly non-trivial, and the automation of the iterative optimization procedure, which can be done using black-box optimization algorithms. \cite{Ozaki2020} 
Initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database. \cite{Belsky2002, GatesRector2019}
However, for novel observed structure types, simply matching to a database of known structures will not yield good results. Therefore, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses.

Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment. \cite{Agrawal2019, Surdu2023}
Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess and running ML-automated Rietveld refinement. \cite{Surdu2023,Feng2019} 
So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020} , machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals. \cite{Schopmans2023}

% Predicting structure information with machine learning
Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
lattice parameters, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021}, and crystallite size from simulated diffractograms.
However, the performance substantially drops off when these models are applied to data originating from experiments [cite 2-3 papers, incl. ours].
The patterns observed in experiments suffer from a host of imperfections that need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns. This includes noise, background, preferred orientation of the sample, ...
Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms.

@@

However, even small labeled and also unlabeled experimental diffractograms hold significant value in practice. First of all, labeled data can be used to better test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset.

>>>>

% On Rietveld refinement
After fabricating a new material, a number of analysis techniques can be used to characterize the sample.
The most important method to determine the crystal structure of a new material is powder X-ray diffraction. 
To analyze pXRD data, an initial guess is optimized using Rietveld refinement. In this process, an initial crystal structure including unit cell parameters is provided by the operator of the analysis software and fitted to the observed diffractogram in an iterative optimization process based on simulated forward prediction of expected pXRD diffractograms given the current crystal structure.\cite{Dinnebier2019, Cano2021}
As Rietveld refinement is a local optimization method, the result of the refinement procedure is generally only as good as the initial guess it was provided with.
%Therefore, it strongly relies on the experience of the operator to provide an accurate initial guess of the crystal structure.

% Manual vs (ML) automated Refinement
Employing manually operated Rietveld refinement, including determining a good initial structure for the optimization, is not without issues.
It is certainly not scalable to the degree required to keep up with advances in throughput and efficiency in other steps of the experimentation pipeline.
Automation of the Rietveld refinement requires the automation of finding initial guess structures, which is highly non-trivial, and the automation of the iterative optimization procedure, which can be done using black-box optimization algorithms. \cite{Ozaki2020} 
Initial guesses for Rietveld refinement are often obtained by matching the observed diffractogram to patterns from a database. Commonly used databases include the "Powder Diffraction File"\cite{GatesRector2019}, the "Inorganic Crystal Structure Database" \cite{Belsky2002} and the "Crystallography Open Database" \cite{Grazulis2009}.
An initial guess obtained from such a database is by no means guaranteed to lead to an accurate structure solution, even if it is iteratively revised, either manually or using blackbox optimization.\cite{Ozaki2020}
Especially for novel observed structure types, simply matching to a database of known structures will not yield good results. Therefore, more complex crystal structure solution approaches need to be applied to find suitable crystal structure guesses.
Machine learning has the potential to speed up the manual analysis of powder diffractograms and keep pace with an automated high-throughput experimentation environment. \cite{Agrawal2019, Surdu2023}
Models can be trained to predict crystal structure information directly given a diffractogram, or models can be used to automate the conventional workflow of finding an initial guess \cite{Surdu2023} and running ML-automated Rietveld refinement. \cite{Feng2019} 
So far, however, due to an absence of labeled datasets with experimental diffractograms\cite{Wang2020}, machine learning in this domain has largely relied on diffractograms simulated from known structures\cite{Park2017, Lee2023} or, most recently, from generated synthetic crystals. \cite{Schopmans2023}

% Predicting structure information with machine learning
Models trained on datasets with simulated diffractograms have already shown very strong performance in predicting
lattice parameters\cite{Dong2021, Chitturi2021, Habershon2004}, spacegroup \cite{Schopmans2023, Oviedo2018, Park2017, Vecsei2018, Zaloga2020, Suzuki2020, Chakraborty2021}, and crystallite size \cite{Dong2021, Chakraborty2021} from simulated diffractograms.
However, the performance substantially drops off when these models are applied to data originating from experiments [cite 2-3 papers, incl. ours].
The patterns observed in experiments suffer from a host of imperfections that need to be modeled accurately in order to be able to transfer models trained on simulated data to experimental patterns. This includes noise, background, preferred orientation of the sample, ...
Most approaches currently model these imperfections as a simple signal based on a heuristic functional form added to the simulated diffractograms.

@@

However, even small labeled and also unlabeled experimental diffractograms hold significant value in practice. First of all, labeled data can be used to better test and benchmark existing and new automated analysis approaches, which yields a better comparison than when using the very limited currently available datasets, such as the commonly used RRUFF mineral dataset. \cite{Armbruster2015}

### 2024-06-09 chapters/03_SOTA_ML.tex
Commited first paragraph in "Our dataset" section from external author

Using simulated diffractograms based on structures from the ICSD, Lee et al. managed to train a deep convolutional neural network (CNN) that is able to classify occurring phases in diffractograms of a specific compound pool. \cite{Lee2020}
Park et al. reached a test accuracy of roughly 81\% for a CNN, which classifies space groups of simulated single-phase diffractograms. \cite{Park2017}
A regression analysis on lattice parameters within a broader framework encompassing all material classes was conducted by Chitturi et al.\cite{Chitturi2021} They developed a distinct CNN for each crystal system, utilizing a merged dataset from both the ICSD and the Cambridge Structural Database, and managed to achieve a mean absolute percentage error of about \SI{10}{\percent} for the lattice lengths, although they encountered difficulties in accurately predicting angles.

>>>>

Using simulated diffractograms based on structures from the ICSD, Lee et al. managed to train a deep convolutional neural network (CNN) that is able to classify occurring phases in diffractograms of a specific compound pool. \cite{Lee2020} Building upon their work they also developed models based on fully convolutional neural networks and transformer encoders, that predict the crystal system and the spacegroup and other structural properties, such as the band gap \cite{Lee2022}. With their best model for the crystal system prediction on ICSD structures, they achieved a test accuracy of \SI{92.2}{\percent}.
Park et al. reached a test accuracy of roughly 81\% for a CNN, which classifies space groups of simulated single-phase diffractograms. \cite{Park2017}
A regression analysis on lattice parameters within a broader framework encompassing all material classes was conducted by Chitturi et al.\cite{Chitturi2021} They developed a distinct CNN for each crystal system, utilizing a merged dataset from both the ICSD and the Cambridge Structural Database, and managed to achieve a mean absolute percentage error of about \SI{10}{\percent} for the lattice lengths, although they encountered difficulties in accurately predicting angles.

### 2024-06-09 chapters/04_our_dataset.tex
Pushed resolved comments
Deleted commented out initial draft material
Included machine learning on Powder Diffraction references
Commited first paragraph in "Our dataset" section from external author

% Notes:
% The subsections of this section should be the 
% individual contributions that we get from labs

% Sharing raw powder x-ray diffraction data: Challenges and benefits https://www.researchgate.net/publication/329256004_Sharing_powder_diffraction_raw_data_Challenges_and_benefits

@@

We are still in the process of growing the dataset and additional contributions are still very welcome. To find out more about how to contribute to this dataset, visit our website specially designed to collect this dataset, https://xrd.aimat.science \\

>>>>

% TODOS:
% - This paper talks about the importance of sharing raw powder X-ray diffraction data. I'd definitely cite it ->  \cite{Aranda2018}

%- The subsections of this section should be the 
% individual contributions that we get from labs

%------------------------------------------$

@@

%% Paragraph by FX Coudert and Arthur Hardiagon
%% !! work in progress !!

We extracted experimental pXRD data from the Crystallography Open Database (COD).\cite{Grazulis2009, Vaitkus2023}  The COD is, to our knowledge, the largest open-access collection of experimental crystal structures of organic, inorganic, metal-organic compounds and minerals, containing more than 500,000 entries.\footnote{Available online at \url{https://www.crystallography.net/cod/}} The data in the COD is placed in the public domain and licensed under the CC0 License. Of the entire COD database, 5432 structures (1\% of the database) contained at least one tag from the {CIF\_POW} dictionary, i.e., relating to powder diffraction studies --- which is expected, since most crystal structures are resolved from single crystal diffraction. From these 5432 files, most contained only metadata related to the powder diffraction experiment, but did not include the raw data of the pattern itself. We could extract raw, experimental pXRD patterns from 1063 files in total, after curation of a small number of files with clearly invalid data.

The pXRD data from the COD database are of high quality, with a median resolution of $\Delta(2\theta) = 0.013$\textdegree and an average number of 9190 points measured per pattern.

% todo : short analysis of the data

%%

We are still in the process of growing the dataset and additional contributions are still very welcome. To find out more about how to contribute to this dataset, visit our website specially designed to collect this dataset, https://xrd.aimat.science .
